---
title: "Activity 09 - Different Confidence Interval Methods"
author: "STS2720 (Dr. VanKrevelen)"
date: 'Updated: 9/29/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
library(moderndive)
```

We've learned three different methods for calculating confidence intervals (percentile, SE, and theory-based). If you're wondering which one you should use, the answer is that it usually doesn't matter much. As we discussed in Notes 9, the three methods all basically...

- are centered at or near your estimate
- use your sample or bootstrap distribution which is based on your sample to estimate a standard error
- and use a theoretical/bootstrap distribution to account for your level of confidence

Let's use the `house_prices` dataset from the `moderndive` package to illustrate how confidence intervals work.

1. Calculate the mean `price` of the houses and call it `mu_price`. We will consider this the population mean. Write the value you get below your code using proper notation. (Note: Normally we won't know this value, but it will help us illustrate how confidence intervals work)

```{r}

```

**ANSWER HERE**

<br>


2. I've included code below to take a sample of 50 houses and stored it in an object called `houses_sample`. Calculate the sample mean price of your sample of houses and store it in `xbar_price`. Write the value you get below your code using proper notation. (Note: Change the seed to a number of your own choosing first.)

```{r}
set.seed(2021929)
houses_sample <- sample_n(house_prices, size = 50)

```

**ANSWER HERE**

<br>


3. If we wanted to make a confidence interval using the theory-based method, we could calculate it with this formula: $\bar{x} \pm t^*_{n-1} * \frac{s}{\sqrt{n}}$ or we can use the `t.test()` function. Let's use R to create a 50% confidence interval and a 90% confidence interval.

```{r}

```

**CIs HERE**

<br>


4. Was the population mean (from question 1) inside your 50% interval? What about your 90% interval?

**ANSWER HERE**

<br>


5. Let's try making the same two intervals (50% and 90%) using the SE method. We will need to create a bootstrap distribution first.

```{r}

```

**CIs HERE**

<br>


6. Compare your intervals with this method to your intervals with the theory-based method. Do they have the same center? Are the endpoints similar? Explain why this is.

**ANSWER HERE**

<br>


7. Use the bootstrap distribution to generate the same two intervals using the percentile method. Do they have the same center as the other two? Are the endpoints similar? Explain what's going on.

```{r}

```

**ANSWER HERE**

<br>


---

Since this is the end of your confidence interval discussion (for now), we will use this space to answer any questions you have or review concepts.