---
title: "Activity 11A - Hypothesis Test for a Proportion"
author: "STS2720 (Dr. VanKrevelen)"
date: 'Updated: 10/16/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
```

The best way to get more comfortable with conducting hypothesis tests is to practice. 

Let's suppose that we're interested in studying the concept of home field advantage in Major League Baseball last season. If there's no such thing as home field advantage, the home team will win just as often as the away team (half the time). If there *is* an advantage to the home team, they will win over half the time. To carry out this test, I will use a significance level of $\alpha = 0.10$.

<br>

1. Step 1 - Write out our null and alternative hypotheses below. Define $p$ in this specific example.

**$H_0:$**

**$H_A:$**

**p = **

<br>


2. Step 2 - I gathered data from two random dates, which you can read into R below. Summarize this data in context of this problem.

```{r}
mlbgames <- data.frame(date = c(rep("8/21/20", 14), rep("9/7/20", 11)),
                       winner = c("H", "A", "A", "A", "H",
                                  "H", "H", "H", "H", "H",
                                  "H", "A", "A", "A",
                                  "H", "A", "A", "A", "H",
                                  "H", "H", "H", "H", "H",
                                  "H"))
```

$\hat{p} =$

<br>


3. Step 3 - Use the `infer` package to create a null distribution and shade in your p-value on the picture. Then calculate the p-value and write a sentence about what the p-value means for this specific example. (Note: You aren't giving a formal conclusion yet.)

```{r, eval = FALSE}
set.seed()
mlb_nulldist <- mlbgames %>%
  specify() %>%
  hypothesize() %H%
  generate() %>%
  calculate()
```

**p-value = **

**Sentence: **

<br>


4. Step 4 - Write a formal conclusion based on your p-value.

**Conclusion: **

<br>


5. Write a couple sentences about why we reached this conclusion. Do you think it's possible we're "wrong"? What could we do to feel more confident in our answer?

**ANSWER**

<br>

---

6. Want extra practice? Try to repeat this process for NBA games last season (before their COVID-19 bubble started). Below is data collected from two random dates last NBA season.

```{r}
nbagames <- data.frame(date = c(rep("1/11/20", 7), rep("2/23/20", 7)),
                       winner = c("H", "H", "A", "A", "H", "H", "H",
                                  "H", "H", "A", "H", "H", "H", "H"))
```

**STEP 1**

<br>


**STEP 2**

<br>


**STEP 3**

<br>


**STEP 4**

