---
title: "Activity 11B - Hypothesis Test for Comparing Proportions"
author: "STS2720 (Dr. VanKrevelen)"
date: 'Updated: 10/16/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
```

# COVID-19 Antibodies Example

The example for this activity comes from a [New York Times article](https://www.nytimes.com/2020/08/19/health/coronavirus-fishing-boat.html) about a ship that had a coronavirus outbreak while at sea. Before setting sail, the crew had been tested for COVID-19 antibodies and 3 of the 120 sailors had these antibodies while the other 117 did not. Below is a quote from the article.

> The three sailors confirmed to have neutralizing antibodies did not test positive for the virus during the course of the study; 103 of the remaining 117 became infected. "These numbers may be small, but they’re highly significant", Dr. Greninger said.

We will conduct a hypothesis test to see if we can see what Dr. Greninger means when he says these results are "highly significant".

<br>


1. Step 1 - How could we set up our hypotheses based on our likely question of interest? Try doing this both in symbols and in words.

**ANSWER**

<br>


2. Step 2 - The data is mentioned at the top of this activity. I've included code to read it into R below. Summarize the data in relation to our question of interest.

```{r}
ship <- data.frame(antibodies = c(rep("yes", 3), rep("no", 117)),
                   infected = c(rep("no", 17), rep("yes", 103)))
```

<br>


3. Step 3 - Create a null distribution and visualize it with a p-value shaded in. Then calculate the p-value and write a sentence about what it means in this context.

```{r}

```

**p-value = **

**Sentence: **

<br>


4. Step 4 - Write a formal conclusion for this example.

**Conclusion:**

<br>


5. Were the results "highly significant"? In other words, are we still able to learn something about the role of antibodies despite only having three people in the group that had them?

**ANSWER**

<br>


6. Besides the sample size, are there any other reasons we may want to be cautious about attributing the whole difference we're seeing to whether the person had antibodies or not?

**ANSWER**

<br>


---

# Extra practice - Hydroxychloroquine Example

Last year a study was conducted looking at the efficacy of hydroxychloroquine in reducing the likelihood of contracting COVID-19. A link to the study is [here](https://www.nejm.org/doi/pdf/10.1056/NEJMoa2016638?articleTools=true). (Note: This example is looking at just one small part of a much larger study.)

1. Step 1 - What hypotheses should researchers use to test whether fewer people will test positive when taking hydroxychloroquine compared to a placebo? Be sure to define your parameters.

<br>


2. Step 2 - In the study, 49 of the 414 subjects taking hydroxychloroquine became infected with COVID-19. 58 of the 407 subjects taking the placebo became infected. Code to read in the data is below. Create a summary of the data that is relevant to this problem. (Bonus: What sort of graph might we want to look at?)

```{r}
study <- data.frame(treatment = c(rep("hydroxy", 414), rep("placebo", 407)),
                    infected = c(rep("yes", 49), rep("no", 414 - 49),
                                 rep("yes", 58), rep("no", 407 - 58)))
```

<br>


3. Step 3 - Use the `infer` package to create a null distribution and to calculate a p-value.

<br>


4. Step 4 - Write a formal conclusion in context of the problem. Use $\alpha = 0.05$.

<br>


5. The authors of the paper say that the study was only designed with sufficient power to find a reduction of at least 50 percent. What do you think this means? What do you think researchers do to make a test "powerful" enough to detect smaller differences?

