---
title: "Activity 09b - Assessing Normality in a Sample"
author: "STS2720 (Dr. VanKrevelen)"
date: 'Updated: 10/6/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r}
# Choose settings for graphs and samples
bins <- 10
n1 <- 200
n2 <- 20
set.seed(1234)
```

In Notes 9, we learned that some theoretical methods require the condition that the data comes from a population with a normal distribution. When we use these methods we *assume* our data comes from a normal distribution, so it's worth exploring how we decide whether this is or is not a reasonable assumption to make.

Let's start by creating four different "populations". The code below will create a data frame called `population` that will contain four different variables, one for each of these populations..

```{r}
population <- data.frame(pop1 = rnorm(10000),
                         pop2 = rt(10000, 5),
                         pop3 = rexp(10000),
                         pop4 = rbeta(10000, 5, 1))
```


1. Let's take a sample of `r n1` observations from each population. We can do this using the `sample_n()` function. Below are graphs of our four samples. Which sample(s) do you think come from normal distributions? How are you deciding?

```{r, eval = FALSE}
sample_200 <- sample_n(population, size = n1)
g1 <- ggplot(sample_200) +
  geom_histogram(aes(x = pop1), bins = bins, color = "white")
g2 <- ggplot(sample_200) +
  geom_histogram(aes(x = pop2), bins = bins, color = "white")
g3 <- ggplot(sample_200) +
  geom_histogram(aes(x = pop3), bins = bins, color = "white")
g4 <- ggplot(sample_200) +
  geom_histogram(aes(x = pop4), bins = bins, color = "white")
grid.arrange(g1, g2, g3, g4, ncol = 2)
```

**ANSWER HERE**

<br>


2. What about if our samples each had `r n2` observations instead of `r n1`? Which sample(s) do you think come from normal distributions? How are you deciding?

```{r, eval = FALSE}
sample_20 <- sample_n(population, size = n2)
g1 <- ggplot(sample_20) +
  geom_histogram(aes(x = pop1), bins = bins, color = "white")
g2 <- ggplot(sample_20) +
  geom_histogram(aes(x = pop2), bins = bins, color = "white")
g3 <- ggplot(sample_20) +
  geom_histogram(aes(x = pop3), bins = bins, color = "white")
g4 <- ggplot(sample_20) +
  geom_histogram(aes(x = pop4), bins = bins, color = "white")
grid.arrange(g1, g2, g3, g4, ncol = 2)
```

**ANSWER HERE**

<br>


3. Another option is to use a graph called a Q-Q Plot. The Q's stand for quantile. What does this word sound like? What do you think it means?

**ANSWER HERE**

<br>


4. A Q-Q plot is like a scatter plot where one axis represents quantiles from our data and the other axis represents quantiles from a normal distribution. Thus, each data point in our sample is plotted against where we would expect that quantile to fall for a normal distribution. If our data is from a normal distribution, we expect the points to fall along a 45 degree line. If not, the points may have a different pattern. Let's try this for our samples of `r n1` observations from our four populations.

```{r, eval = FALSE}
g1 <- ggplot(sample_200, aes(sample = pop1)) +
  geom_qq() +
  geom_qq_line()
g2 <- ggplot(sample_200, aes(sample = pop2)) +
  geom_qq() +
  geom_qq_line()
g3 <- ggplot(sample_200, aes(sample = pop3)) +
  geom_qq() +
  geom_qq_line()
g4 <- ggplot(sample_200, aes(sample = pop4)) +
  geom_qq() +
  geom_qq_line()
grid.arrange(g1, g2, g3, g4, ncol = 2)
```

Which of these graph(s) looks *most* like the points fall along a 45 degree line? Which of these graph(s) seem to clearly have a different pattern?

**ANSWER HERE**

<br>


5. The histogram seemed to work OK for bigger sample sizes too. Let's see if the QQ plot works better for smaller samples. We will make a QQ plot for each of our samples of size `r n2`.

```{r, eval = FALSE}
g1 <- ggplot(sample_20, aes(sample = pop1)) +
  geom_qq() +
  geom_qq_line()
g2 <- ggplot(sample_20, aes(sample = pop2)) +
  geom_qq() +
  geom_qq_line()
g3 <- ggplot(sample_20, aes(sample = pop3)) +
  geom_qq() +
  geom_qq_line()
g4 <- ggplot(sample_20, aes(sample = pop4)) +
  geom_qq() +
  geom_qq_line()
grid.arrange(g1, g2, g3, g4, ncol = 2)
```

Which of these graph(s) looks *most* like the points fall along a 45 degree line? Which of these graph(s) seem to clearly have a different pattern?

**ANSWER HERE**

<br>

6. I've set up this document so that you can change the two sample sizes we look at (and the number of bins for the histograms). Try playing around with the different settings a bit. What do you notice?

**ANSWER HERE**

<br>

Note: You can explore these ideas more with [this app](https://vank-stats.shinyapps.io/samplehistograms/) as well. The app was made using the Shiny package.