---
title: "Activity 05 - Sampling from a Population"
author: "STS2720 (Dr. VanKrevelen)"
date: 'Updated: 9/9/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will use the following two R packages for this activity.

```{r, warning = FALSE, message = FALSE}
library(ggplot2)
library(moderndive)
library(dplyr)
```

Our focus for the upcoming sets of notes has to do with using a sample to make informed conclusions about a population. To be able to do this well, we need to have a good understanding of what happens when we randomly sample from a population.

The `diamonds` dataset in the `ggplot2` package contains 53,940 diamonds. Let's pretend that this is our population and that we don't actually have access to the whole population. Instead, we will take a sample from the population to estimate the population mean diamond price.

1. First as some review... what symbol do we use to represent a population mean? What about for the sample mean?

Population mean: $\mu$

Sample mean: $\bar{x}$

<br>

2. When we take random samples using R, it may be helpful for us to do so in a way that someone else can recreate our exact results. To do this, we set the "seed" where the randomization process will start. Replace the number inside the `set.seed()` function with a number of your choosing. Then use the `rep_sample_n()` function to take a random sample of 50 diamonds.

```{r}
set.seed(210913)
diamonds_50 <- rep_sample_n(diamonds, size = 50)
```


<br>

3. Calculate your sample mean price and we will compare numbers across everyone in our class.

```{r}
mean(diamonds_50$price)
```


<br>

4. Now let's take a sample of 500 diamonds and calculate the sample mean of this new sample. Again we will compare numbers across the class. Run `set.seed()` at the top of this code too with a new seed chosen. You can base your code below on the code from Question 2.

```{r}
set.seed(759367)
diamonds_500 <- rep_sample_n(diamonds, size = 500)
mean(diamonds_500$price)
```


<br>

5. What do you notice about the sample means we got with samples of 50 diamonds vs. with samples of 500 diamonds?

Both of the groups of sample means are centered around 3800 or so but the samples of 50 diamonds range from means in the upper 2000s to mid 4000s. The means of samples of 500 diamonds ranged from the 3700s to right around 4000.

<br>

6. Our class only has 10 people (9 students plus Dr. V). It'd be better if we could look at A LOT of random samples. We can do this with the `rep_sample_n()` function using the `reps = __` argument. Let's take 1,000 samples of 50 diamonds, calculate the sample mean for each one, and then make a histogram of our results. We can do this in multiple steps or we can use piping to do it all at once.

```{r}
my1000samps <- rep_sample_n(diamonds, size = 50, reps = 1000) %>%
  group_by(replicate) %>%
  summarize(xbar = mean(price))

ggplot(my1000samps) +
  geom_histogram(aes(x = xbar), color = "orange") +
  geom_vline(xintercept = 3932.8, color = "lightblue", size = 2)
```


<br>

7. Let's do the same thing as above but with 500 diamonds in each of our 1,000 samples.

```{r}
my1000samps_500 <- rep_sample_n(diamonds, size = 500, reps = 1000) %>%
  group_by(replicate) %>%
  summarize(xbar = mean(price))

ggplot(my1000samps_500) +
  geom_histogram(aes(x = xbar), color = "orange") +
  geom_vline(xintercept = 3932.8, color = "lightblue", size = 2)
```


<br>

8. What do you notice when you compare the two histograms? What is similar about them? What is different?

Whether I'm taking samples of 50 or 500 diamonds, the graph of sample means is centered a little bit under $4,000. However, the spread of the sample means is much wider for samples of 50 diamonds than for samples of 500 diamonds.

<br>

9. Normally we don't know the population mean (that's why we estimate it), but in this case I can calculate it from the `diamonds` data set. The population mean price is $3,932.80. Where does this number show up in the two graphs you made? How could we use this result to talk about the **standard error** in the two graphs?

We are taking random samples from the population, so the samples should sort of reflect the population. Sometimes our sample means will be above the population mean and sometimes they will be below, but these should balance out and average the population mean in the middle of our graph.

The standard error is the standard deviation of my sample means (this *different* than the standard deviation of the population or sample itself). In other words, the standard error tells us how spread out our estimates of the population will be if we took a whole bunch of random samples. The "error" refers to how far our estimates (i.e. statistics) are from the parameter of interest.

We can see that when our samples are bigger there is less "error" between the sample means and the population mean. In other words, the standard error is smaller for large sample sizes because our sample means tend to be closer to the population mean.

<br>

10. What do you think your answers to the previous two questions can tell you about random sampling and estimating parameters with statistics? We can explore this idea more with [this link](http://onlinestatbook.com/stat_sim/sampling_dist/).

