---
title: "Activity 13 - ANOVA Tests for Many Means"
author: "STS2720 (Dr. VanKrevelen)"
date: 'Updated: 10/26/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
```


# Example - Fidget spinners

In the fidget spinner example at the end of our last activity, I removed the group that was actively using the fidget spinner. However, that group was likely a key part of the experiment. Now that we've learned above ANOVA tests, let's conduct a hypothesis test to compare the population mean memory scores of *all three groups*. We will use $\alpha = 0.1$ for our hypothesis test.


1. **Step 1** - Write out our hypotheses

Write out the appropriate hypotheses to conduct an ANOVA hypothesis test for this problem.

$H_0: \mu_1 = \mu_2 = \mu_3$

$H_A:$ at least one population mean memory scores is different from another


<br>

2. **Step 2** - Gather and summarize data

There is code below to read the data in from our course GitHub repository. Generate an appropriate graph to compare the `MemoryScore`'s for the three groups.

```{r}
fidget <- read.csv("https://raw.githubusercontent.com/vank-stats/STS2720-Fall2021/main/fidget_spinners.csv")

ggplot(fidget) +
  geom_boxplot(aes(x = MemoryScore, y = Condition))
```

Use the `infer` package to calculate an F statistic for this data. What does the value we calculate mean in this context?

```{r}
fidget_F <- fidget %>%
  specify(formula = MemoryScore ~ Condition) %>%
  calculate(stat = "F")
fidget_F
```

$F = 4.51$

**Interpretation of F: The average variability between the conditions is 4.51 times as big as the average variability within the groups.**


<br>

3. **Step 3** - Compare the data to $H_0$

Use the `infer` package to generate a null distribution of F values. Visualize your p-value on the distribution.

```{r}
set.seed(1031)
fidget_null <- fidget %>%
  specify(formula = MemoryScore ~ Condition) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F")

visualize(fidget_null) +
  shade_pvalue(obs_stat = fidget_F, direction = "right")
```

Calculate the p-value and write a sentence explaining what it means in context.

```{r}
fidget_null %>%
  get_pvalue(obs_stat = fidget_F, direction = "right")
```

$p-value = 0.018$

**P-value Interpretation: In a world where memory score is independent of which group someone is in (the population mean scores are all the same), there is only a 1.8% of getting an F statistic at least as large as we saw.**



<br>

4. **Step 4** - Reach a conclusion

**Conclusion: The p-value of 0.018 is under our significance level, so we can reject the null hypothesis. We can be very certain that the alternative hypothesis is true and at least one condition has a different population mean memory score than another.**



<br>

---

# Simultaneous Confidence Intervals

If we concluded that there is a difference between at least some of the population means, our next step should be to determine *which* groups are different. We'll do this with confidence intervals, but there's a catch...

- Before getting into confidence intervals. Let's suppose I plan to flip a fair coin. How confident can I be that it will come up heads?

**I'm 50% confident I will get heads**

<br>

- If I flip two coins, how confident can I be that *both* will come up heads? What about with three coins?

**Two coins: $0.5 * 0.5 = 0.25* --> I am 25% confident that both come up heads**

**Three coins: $0.5 * 0.5 * 0.5 = 0.125* --> I am 12.5% confident that all three come up heads**

<br>

- We can compare this to confidence intervals. If I make two 90% confidence intervals (for example), I can't be 90% confident that they are *both* correct. How confident am I? What about with three 90% confidence intervals?

**Two 90% intervals --> .9 * .9 = .81 (or 81% confident both contain the true difference in population means)**

<br>



### Bonferroni's Adjustment

For our ANOVA test we did, we will need to make three confidence intervals. (Group 1 - Group 2, Group 1 - Group 3, and Group 2 - Group 3). In general, we can talk about confidence intervals as $(1-\alpha)*100\%$ intervals. Something called **Bonferroni's method** can be used to help us choose a new $\alpha$ that will make it so we can be 90% confident *overall*.

$$\text{Bonferroni's method}: \alpha^* = \frac{\alpha}{m}$$

where $\alpha$ is our overall significance level and $m$ is the number of intervals we will make.

(Note: While it's called Bonferroni's method, it was statistician [Olive Jean Dunn](https://en.wikipedia.org/wiki/Olive_Jean_Dunn) who first discovered the application to confidence intervals.)

<br>

5. What will $\alpha^*$ be if we make three intervals and want to be 90% confident overall? How confident will we be in the individual intervals?

```{r, eval = TRUE}
alphastar <- .1 / 3
```


<br>

6. Verify that we can be 90% confident in all three intervals together.

```{r}
(1 - alphastar)^3
```


<br>

Unfortunately, there isn't a convenient simultaneous CI method in the `infer` package using simulation (or built into R). We can use the `DescTools` package to generate theory-based versions of these CIs though. Install this package in your R console then run the code below.

```{r, eval = TRUE}
library(DescTools)
fidget_anova <- aov(MemoryScore ~ Condition, data = fidget)
PostHocTest(fidget_anova, conf.level = 0.9, method = "bonferroni")
```

7. Which group(s) can we be 90% confident have different population means from one another? Justify your answer.

**We can be 90% confident spinner and no spinner have different population means than one another because the interval comparing the two is entirely negative.**

<br>


8. Can we be 90% confident one group has the highest population mean? What about the lowest? Justify your answer.

**No. We can be 90% confident the population mean memory score is higher for the no spinner group than the spinner group, but when we compare either group to the spinner present group, the interval contains both positive and negative numbers.**

<br>


9. What might our overall conclusion look like? Try writing one out that incorporates the information that we've learned from the fidget dataset.

**We are 90% confident that the population mean memory score is higher for people with no spinner present than for those actively using a spinner. We are not 90% confident whether the population mean memory score for people with a spinner nearby is higher or lower than either of the other two population means.**

